Labor Markets and Technological Change: Evidence from Electronic Health Records

### 2022-06-14 Paper/Paper.tex
Updates from Overleaf

An institutional assumption that I make is that physicians working in EHR-adopted hospitals are constrained to use the technology. My outcomes have two key determinants: the choice of physicians to learn and utilize the technology, and the realized effect of electronic health records themselves. The point of this paper is to identify the second determinant, and thus I assume that when a physician continues to work in an electronic record utilizing-hospital, they utilize the electronic health record system fully. That is, there are no physicians who stay actively working in a hospital but choose not to partake in the electronic health record used by the hospital. It is reasonable to believe that physicians are fully utilizing the EHRs in the hospital they work in, since by 2015 electronic health records were widely implemented and considered unavoidable. An objection to this assumption may be that physicians stay in a hospital, but the hospital hires an employee to do technology work for the physician. In this case, the effects seen could be from having the data assistant instead of EHR use. I investigate this in Section \ref{sec:dataass}.

@@

@@

@@

However, for both patient and claim count, the results include physicians who shift place of work, which may lead to biased estimates since a physician could switch hospitals due to an EHR and then change practice behavior due to the move, not due to the EHR. Thus, I limit the sample to only physicians who work with the same one hospital through the entire sample. These physicians do not retire or see patients in another hospital. Thus, when an EHR is implemented, the physician remains utilizing the EHR for the remainder of the period. I present the results for this limited sample in Figure \ref{fig:limitedsample}. These results are similar to those presented for the full sample, suggesting that physicians become more productive in terms of patient count in the first two years after implementation.

@@

\section{Physician EHR Avoidance}

\subsection{Physician Influence of EHR Decision in Hospitals}

@@

\begin{figure}[p]
    \centering
    \caption{Effect of EHR Exposure on Patient Count by Group}
    \includegraphics[scale=.35]{Objects/patient_group.pdf}
    \label{fig:patientgroup}
\end{figure}

\begin{figure}[p]
    \centering
    \caption{Effect of EHR Exposure on Claim Count, Limited Sample}
    \includegraphics[scale=.65]{Objects/claim_plot.pdf}
    \label{fig:claim}
\end{figure}

\begin{figure}[p]
    \centering
    \caption{Effect of EHR Exposure on Patient Count, Limited Sample}
    \includegraphics[scale=.45]{Objects/limitedsample_plot.pdf}
    \label{fig:limitedsample}
\end{figure}

\begin{figure}[p]
\centering
\caption{Frequency of Data Assistant Enumeration by Year}
\includegraphics[scale=.5]{Objects/dataassistant_histogram.pdf}
\label{fig:dataassistant_histogram}
\end{figure}

>>>>

An institutional assumption that I make is that physicians working in EHR-adopted hospitals are constrained to use the technology. My outcomes have two key determinants: the choice of physicians to learn and utilize the technology, and the realized effect of electronic health records themselves. The point of this paper is to identify the second determinant, and thus I assume that when a physician continues to work in an electronic record utilizing-hospital, they utilize the electronic health record system fully. That is, there are no physicians who stay actively working in a hospital but choose not to partake in the electronic health record used by the hospital. It is reasonable to believe that physicians are fully utilizing the EHRs in the hospital they work in, since by 2015 electronic health records were widely implemented and considered unavoidable. An objection to this assumption may be that physicians stay in a hospital, but the hospital hires an employee to do technology work for the physician. In this case, the effects seen could be from having the data assistant instead of EHR use. Thus, I investigate hteir existence further.

\begin{figure}[ht]
\centering
\captionsetup{width=.5\linewidth}
\caption{Frequency of Data Assistant Enumeration by Year}
\includegraphics[scale=.5]{Objects/dataassistant_histogram.pdf}
\label{fig:dataassistant_histogram}
\vspace{2mm}
    \caption*{\footnotesize{\textit{Notes:} Histtogram uses NPPES data on the enumeration of certain types of NPIs to estimate the number of new data assistants being hired in each year of my sample.}}
\end{figure}

@@

For both patient and claim count, if the sample includes physicians who shift place of work, this may lead to biased estimates since a physician could switch hospitals due to an EHR and then change practice behavior due to the move, not due to the EHR. Thus, I limit the sample to only physicians who work with the same one hospital through the entire sample. These physicians do not retire or see patients in another hospital. Thus, when an EHR is implemented, the physician remains utilizing the EHR for the remainder of the period.

@@

\begin{figure}[ht]
    \centering
    \caption{Effect of EHR Exposure on Patient Count by Group}
    \includegraphics[scale=.47]{Objects/patient_group.pdf}
    \label{fig:patientgroup}
    \vspace{2mm}
    \caption*{\footnotesize{\textit{Notes:} Plot shows average group time treatment effects for each treatment group.}}
\end{figure}

@@

\begin{figure}[ht]
    \centering
    \caption{Effect of EHR Exposure on Claim Count}
    \includegraphics[scale=.65]{Objects/claim_plot.pdf}
    \label{fig:claim}
    \vspace{2mm}
    \caption*{\footnotesize{\textit{Notes:} The top panel shows average group time treatment effects aggregated over groups to an event study plot. The bottom show these results for different subgroups of physicians by age. The p-value listed for each graph corresponds to a Wald test for pre-trends. Confidence intervals shown are simultaneous confidence bands accounting for multiple hypothesis testing. Overall ATT for all, $<$ 60, $>=$ 60 with SE in parentheses: 4.04 (33.89), 25.77 (31.26), -161.06 (133.35), respectively.}}
\end{figure}

@@

\section{Physician Influence of EHR Decision in Hospitals}

@@
